---
title: "MDSD 6306 Live Session Week 4 Assignment"
author: "Jenna Ford"
date: "6/4/2019"
output: html_document
---

## Prepwork for the Assignment - Libraries and Downloading Data
```{r prep}

library(downloader)
setwd("C:/Users/b007224/Documents/masters_in_data_science/SMU-MSDS/DDS/Homework/Week5")

download("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit 5/yob2015.txt",destfile="yob2015.txt")
download("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit 5/yob2016.txt",destfile="yob2016.txt")
```

## Question 1 - Data Munging

```{r question1}
#Read in the yob2015.txt file
df <- read.table("yob2016.txt",sep = ";", header=FALSE)

#Assign names to the columns
names(df) <- c("firstName","gender","numberChildren")

#Print summary information for the dataframe
summary(df)

#Print structure information for the dataframe
str(df)

#Identify the mispelled record
grep("yyy$",df$firstName,value=TRUE)

#Remove the mispelled record from the dataframe
y2016 <- df[!(df$firstName==grep("yyy$",df$firstName,value=TRUE)),]

```

## Question 2 - Data Merging

```{r}
#Read in the yob2015.txt file
y2015 <- read.table("yob2015.txt",sep = ",", header=FALSE)

#Assign names to the columns
names(y2015) <- c("firstName","gender","numberChildren")

#Print the last 10 records of the dataset
tail(y2015,10)
```

The last 10 names in this dataset are all male names beginning with 'Z' with a fequency of 5. Sorting is by gender, decreasing frequency and the first letter of the first name. It appears this dataset only includes names that are used 5 or more times.  It's surprising that I have I have never heard of any of these names, considering they are on a popular list of names. But perhaps it's not that surprising considering how infrequent they are. I guess if you want a name that ALMOST no-one else will have, choose one of these names...You could even start a trend!

```{r}
#Merge the datasets together using an inner join
#The instructions only said to merge by Name, but I thought it was more appropriate to merge on name and gender
final <- merge(x=y2016,y=y2015,union("firstName","gender"))

#Rename the columns to be more clear
names(final) <- c("firstName","gender","numberChildren.2016","numberChildren.2015")
```

## Question 3 - Data Summary

```{r}
#Create a total column for the total number of children over the 2 years by name and gender
final$Total <- final$numberChildren.2016 + final$numberChildren.2015

#Total number of people given popular names over the 2 years
sum(final$Total)
```

I want to point out that this isn't the total number of people given popular names, only those where the name was popular in both 2015 and 2016.

```{r}
#Order the dataset by descending Total
final <- final[order(final$Total,decreasing=TRUE),]

#Print out the top 10 most popular names
head(final,10)

#Filter out the male names
final_girls <- final[!(final$gender=="M"),]

#Print the top 10 most popular girl names
head(final_girls,10)

#Write the top 10 most popular girl names to a CSV file, only including the name column
write.csv(final_girls[1:10,1,drop=FALSE],file="top_10_girls.csv",row.names=FALSE)
```

The materials for this homework assignment can be found  [here](https://github.com/JennaFord/SMU-MSDS/tree/master/DDS/Homework/Week5).