---
title: "MDSD 6306 Live Session Week 4 Assignment"
author: "Jenna Ford"
date: "5/29/2019"
output: html_document
---

## Harry Potter Cast

Scrape the [IMDB site](https://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1) to get the full cast for the Harry Potter and teh Dealthy Hallows Part 2 movie. Clean up the dataframe and print out the first 15 records.

```{r harry_potter}
#Load packages
library(stringr)
library(xml2)
library(rvest)
library(tidyr)

#Set URL
url <- "https://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1"

#Read the 3rd table tag
cast_list <- url %>% read_html() %>% html_nodes(xpath='//*[@id="fullcredits_content"]/table[3]') %>% html_table()

#Convert to dataframe
cast_list <- cast_list[[1]]

#Remove the first row and the 1st and 3rd columns, these are all blanks
cast_list1 <- cast_list[-1,-c(1,3)]

#Add column names
names(cast_list1) <- c("Actor","Character")

#Remove extra characters from the dataframe
cast_list2 <- as.data.frame(apply(cast_list1,2,function(x)gsub('\\s+',' ',x)))

#Remove row 93 dynamically
cast_list3 <- cast_list2[!(cast_list2$Actor=="Rest of cast listed alphabetically:"),]

#Split the Actor's fullname into first and last, with middle being inluded with first
cast_list4 <- cast_list3 %>% separate(Actor, into=c("FirstName", "Surname"), sep="[ ](?=[^ ]+$)")

#Print the first 15 rows of the dataframe, I printed enough to show how the middle name was handled
head(cast_list4,15)
```

## San Antonio Spurs Shooting Statistics

Scrape the [ESPN site](http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs) to get stats for the San Antonio Spurs basketball team. Clean up the dataframe and create a bar chart with columns colored by position.

```{r spurs}
#Read in the data from the url
spurs<-read_html("http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs")

#Pull out the tables 
spurs_table<-html_nodes(spurs,"table")
tables <-html_table(spurs_table,fill=TRUE)

# Get the right tables
names <- data.frame(tables[6])
stats <- data.frame(tables[8])

#Combine the tables into one dataframe
df <- cbind(names,stats)

#Remove the 'Total' row
df1 <- df[!(df$Name=="Total"),]

#Separate the 'Name' column into 'Name' and 'Position' columns
df2 <- df1 %>% separate(Name, into=c("Name", "Position"), sep="[ ](?=[^ ]+$)")

#Change the variable type of 'Position' from character to factor (to group for the legend)
df2$Position <- as.factor(df2$Position)

#Change column names to indicate which are percentages
names(df2) <- gsub(x=names(df2), pattern = "\\.$",replacement = "_Perc")

#Verify that numeric columns are properly classified
str(df2)

#Identify Spurs colors (both current and previous colors)
spurscols <- c("#8A8D8F","#000000","#EF426F","#00B2A9","#FF8200")

#Create the bar chart with a legend
par(mar=c(4,12,4,4),mgp=c(10,1,0),las=2)
barplot(df2$FG_Perc
        ,names=df2$Name
        ,horiz=T
        ,xlab=""
        ,ylab="Player Name"
        ,main="Field Goal Percentage Per Game"
        ,col=spurscols[df2$Position]
        ,xlim=range(pretty(c(0,df2$FG_Perc)))
        ,ylim=c(0,17)
        ,legend=TRUE)
legend(x="top",legend=levels(df2$Position),fil=spurscols,bty="n",ncol=3)
title(xlab="Field Goal Percentage",mgp=c(2,1,0))
```

The materials for this homework assignment can be found  [here](https://github.com/JennaFord/SMU-MSDS/tree/master/DDS/Homework/Week4).