---
title: "Craft Beer & Brewery Data Analysis"
author: "Jenna Ford & Reannan McDaniel"
date: "6/29/2019"
output: 
  html_document:
    keep_md: yes
---

## Introduction

The data provided for this analysis consists of 2410 US craft beers from 558 US breweries. Our data analysis that follows highlights the following:

* A look at the data
* How many breweries are in each state
* Missing values for ABV & IBU measurements
* Descriptive information for ABV
* Median ABV & IBU by state
* Beer with the highest ABV and beer with the highest IBU
* Correlation between ABV and IBU

## Setup: Libraries and Reading in Datasets
```{r message=FALSE}
library(RCurl)
library(plyr)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(car)
library(naniar)
library(viridis)
library(scales)

setwd("./Data")

Beers <- read.csv(file="Beers.csv")
Breweries <- read.csv(file="Breweries.csv")

opts <- options()
options(ggplot2.continuous.colour="viridis")
options(ggplot2.continuous.fill="viridis")
```

## A Look at the Data
```{r question2}
BreweriesFullDataset <- merge(Breweries,Beers, by.x = "Brew_ID", by.y = "Brewery_id",all=TRUE)
names(BreweriesFullDataset) <- c("Brew_ID","Brewery","City","State","Beer","Beer_ID","ABV","IBU","Style","Ounces")
head(BreweriesFullDataset,6)
tail(BreweriesFullDataset,6)
```

## Brewery Counts by State
```{r question1}
BreweriesByState <- plyr::count(Breweries$State)

ggplot(BreweriesByState,aes(x=reorder(x,-freq),y=freq,fill=freq)) +
  geom_bar(stat="identity",position="dodge",show.legend=FALSE) +
  xlab("State") +
  ylab("Brewery Count") +
  ggtitle("Brewery Counts by State") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust=0.5))
```

## Missing Values
```{r question3}
gg_miss_var(BreweriesFullDataset) +
  ggtitle("Missing Values by Variable") + 
  theme(plot.title = element_text(hjust = 0.5))
gg_miss_fct(x=select(BreweriesFullDataset,State,ABV,IBU),fct=State) +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust=0.5)) +
  ggtitle("% Missing ABV & IBU Values by State")
```

## Summary Statistics for ABV
```{r question6}
summary(BreweriesFullDataset$ABV)
```

## Median ABV & IBU by State
```{r question4}
Median.ABV.IBU <- BreweriesFullDataset %>% group_by(State)%>% summarise(Median_ABV=median(ABV,na.rm=TRUE),Median_IBU=median(IBU,na.rm=TRUE))

ggplot(Median.ABV.IBU,aes(x=reorder(State,-Median_ABV),y=Median_ABV,fill=Median_ABV)) +
  geom_bar(stat="identity",position="dodge",show.legend=FALSE) + xlab("State") + ylab("ABV") +
  ggtitle("Median ABV by State") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust=0.5))

ggplot(Median.ABV.IBU,aes(x=reorder(State,-Median_IBU),y=Median_IBU,fill=Median_IBU)) +
  geom_bar(stat="identity",position="dodge",show.legend=FALSE) +
  xlab("State") + ylab("Median IBU by State") +
  ggtitle("IBU by State") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust=0.5))
```

## Max ABV & IBU
```{r question5}
BreweriesFullDataset[which.max(BreweriesFullDataset$ABV),]
BreweriesFullDataset[which.max(BreweriesFullDataset$IBU),]
```

## Relationship between ABV & IBU
```{r question7}
scatterplot(BreweriesFullDataset$IBU,BreweriesFullDataset$ABV,xlab="IBU",ylab="ABV",main="Relationship between ABV and IBU",smooth = FALSE, grid = FALSE, frame = FALSE, col="#93e45c")

cor(BreweriesFullDataset$IBU,BreweriesFullDataset$ABV,method=c("pearson","kendall","spearman"),use="complete.obs")
lm <- lm(IBU~ABV,data=BreweriesFullDataset)
summary(lm)$r.squared
```
