---
title: "DDS Live Session Week 12 Assignment"
author: "Jenna Ford"
date: "7/26/2019"
output: html_document
---

## Loading Libraries
```{r message=FALSE}
library(fpp2)
library(RCurl)
library(dygraphs)
library(xts)
library(lubridate)

```

## Warm-Up: Brief Financial Data

The data in this section is for the DAX stock market index from 1991-1999 (from the EuStockMarkets dataset in R). A plot of the data is provided first, followed by a decomposition plot that separates out the seasonal, trend and random components with a multiplicative model.

```{r warmup}
dax = EuStockMarkets[,"DAX"]
plot(dax,ylab = "Closing Price", xlab = "Date", main="DAX Index Daily Closing Prices",col="blue", sub="Market Crash on October 27, 1997")

#I estimated where the market crash of Oct 27, 1997 is on the plot
abline(v=1997.823, col="red")

decomp <- decompose(dax, type="mult")
plot(decomp,col="blue", xlab = "Date",sub="Market Crash on October 27, 1997")
abline(v=1997.823, col="red")
```

## Temperature Data

The data for this section is from the maxtemps dataset in R. It has the maximum annual temperatures for the Moorabbin Airport in Melbourne for 1971-2016. 2 models will be fitted: a simple exponential smoothing model and a damped Holt's linear trend model.The 2 models will be compared to see which is the better fit.

```{r temp}
#Get data for 1990 onward
temp = window(maxtemp, start = 1990)

#SES Model
fit1 = ses(temp, h = 5) 
plot(temp,ylab = "Max Temp", xlab = "Year", type = "o", xlim=c(1990,2021), main="Max Annual Temperatures at Moorabbin Airport")
lines(fitted(fit1), col = "blue", type = "o")
lines(fit1$mean, col = "blue", type = "o")

#Damped Holt's Linear Model
fit2h = holt(temp, alpha = .8, beta = .2, damped = TRUE, initial="optimal", h = 5)
plot(temp,ylab = "Max Temp", xlab = "Year", type = "o", xlim=c(1990,2021), main="Max Annual Temperatures at Moorabbin Airport")
lines(fitted(fit2h), col = "blue", type= "o")
lines(fit2h$mean,col = "blue", type= "o")

#Plot together
plot(temp,ylab = "Max Temp", xlab = "Year", type = "o", xlim=c(1990,2021), main="Max Annual Temperatures at Moorabbin Airport")
lines(fitted(fit1), col = "blue", type = "o")
lines(fit1$mean, col = "blue", type = "o")
lines(fitted(fit2h), col = "red", type= "o")
lines(fit2h$mean,col = "red", type= "o")

#Compare ASE for both models
ASE1 <- sum((fitted(fit1) - maxtemp)^2)/length(temp)
ASE2h <- sum((fitted(fit2h) - maxtemp)^2)/length(temp)

ASE1
ASE2h
```

The ASE for the Simple exponential smoothing model was 5.39 compared to 8.95 for Holt's damped linear model. As such, The SES model performs better using ASE.

```{r temp2}
fit1$model
fit2h$model
```

The AIC and BIC scores were both minimized in the SES model. The SES model performs better using AIC and BIC criterion. Both ASE and AIC/BIC show that SES is the better model.

## The Wands Choose the Wizard

The data for this section is wands sold by Ollivander and Gregorovitch from the Harry Potter Series. First, the 2 data sets will be combined into one timestamp series. Then a dynamic graph will be produced.

```{r wizard}
URL <- getURL("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%2012/Unit12TimeSeries_Ollivander.csv")
ollivander <- read.csv(text=URL, header=FALSE)
names(ollivander) <- c("date","ollivander")

URL <- getURL("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%2012/Unit12TimeSeries_Gregorovitch.csv")
gregorovitch <- read.csv(text=URL, header=FALSE)
names(gregorovitch) <- c("date","gregorovitch")

ollivander$date2 <- mdy(ollivander$date) 
tsollivander <- xts(ollivander$ollivander, ollivander$date2)
names(tsollivander) <- c("ollivander")

gregorovitch$date2 <- mdy(gregorovitch$date) 
tsgregorovitch <- xts(gregorovitch$gregorovitch, gregorovitch$date2)
names(tsgregorovitch) <- c("gregorovitch")

combined = merge(tsollivander,tsgregorovitch)
dygraph(combined, main="Ollivander and Gregorovitch Yearly Wands Sales", xlab="Year", ylab="Wands Sold") %>% dyRangeSelector(height = 100) %>% dySeries("ollivander",color="orange") %>% dySeries("gregorovitch",color="purple") %>% dyShading(from="1995-01-01", to="1999-01-01", color="lightgray") %>% dyHighlight(highlightSeriesOpts=list(strokeWidth=3))
 ```