---
title: "Live Session Unit 3 Assignment"
author: "Jenna Ford"
date: "5/24/2019"
output: html_document
---

## Question 1 - GitHub Cloning
Clone https://github.com/caesar0301/awesome-public-datasets to your local machine.
$ git clone https://github.com/caesar0301/awesome-public-datasets

## Question 2 - Data Summary

```{r getdata}
df <- read.csv("C:/Users/b007224/Documents/masters_in_data_science/doing_data_science/awesome-public-datasets/Datasets/titanic.csv/titanic.csv",header=TRUE)
```

Output the respective counts of females and males aboard the Titanic.

```{r gender}
table(df$Sex)
```

Plot the Frequency of females and males.

```{r barplot}
barplot(table(df$Sex)
        ,xlab="Sex"
        ,ylab='Frequency'
        ,main="Count of Males and Females on the Titanic")
```

Use one apply function to putput the means of Age, Fare and Survival.

```{r means}
sapply(df[,c(2,6,10)],mean,na.rm=TRUE)
```

## Question 3 - Function Building

Download the sleep_data_01.csv dataset from http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/index.html:

```{r sleepdata}
library(repmis)
UrlAddress <- "http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/sleep_data_01.csv"
sleep <- repmis::source_data(UrlAddress)
```
Create a function

```{r func}
sleep.df <- function(dataset){
  #Median age
  median_age <- median(dataset$Age,na.rm=TRUE)

  #Minimum duration of sleep
  min_sleep_dir <- min(dataset$Duration,na.rm=TRUE)
 
  #Maximum duration of sleep
  max_sleep_dur <- max(dataset$Duration,na.rm=TRUE)

  #Mean of the Rosenberg Self Esteem scale (RSES)
  mean_RSES <- mean(dataset$RSES,na.rm=TRUE)
 
  #Standard deviation of the Rosenberg Self Esteem scale (RSES)
  sd_RSES <- sd(dataset$RSES,na.rm=TRUE)

  #Create a data.frame
  report <- data.frame(median_age,mean_RSES/5,sd_RSES/5,max_sleep_dur-min_sleep_dir)
 
  #Rename the colmns
  colnames(report) <-c("MedianAge","SelfEsteem","SE_SD","DurationRange")
  
  #Round the report to 2 digits
  round(report,2)
}
```

Call the sleep.df function

```{r callfunc}
sleep.df(sleep)
```

## Question 4 - FiveThirtyEight Data

```{r five38}
library(fivethirtyeight)
df <- college_recent_grads
```

Check out [The Economic Quide to Picking a College Major](http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/)
for the news story from FiveThirttyEight.

```{r five38dim}
#Dimension of the data frame
dim(df)

#Columns names of the data frame
names(df)
```

## Question 5 - Data Summary for FiveThirtyEight Data

```{r summary}
#Column names of the data frame
names(df)

# Number of columns
ncol(df)
```

Generate a count of each unique major_category in the data frame.

```{r major}
library(plyr)
major_count <- ddply(df, ~ major_category,summarize,n_uniq_major = length(unique(major)))
class(major_count)
```

Plot the count of majors by major category.I couldn't get all the labels and axis labels to show up on the graph in only one barplot statement. I used the title statment to get the x axis label visible on the plot.

```{r barplot2}
par(mar=c(4,14,4,4),mgp=c(12,1,0),las=2)
barplot(major_count$n_uniq_major
        ,names.arg = major_count$major_category
        ,xlab=""
        ,ylab="Major Categories"
        ,main="Counts of Majors by Categories"
        ,cex.names=.75
        ,horiz=T
        ,col="blue"
        ,xlim=range(pretty(c(0,major_count$n_uniq_major))))
title(xlab="Majors",mgp=c(2,1,0))
```

Write out csv file.
```{r csv}
write.csv(college_recent_grads,file="college_recent_grad.csv")
```

## Question 6

My new repository is [here](https://github.com/JennaFord/SMU-MSDS).

The materials for this homework assignment are [here](https://github.com/JennaFord/SMU-MSDS/tree/master/DDS/Homework/Week3).